Supporting Information - Tackling discrepancies in freshwater invertebrate trait databases: Harmonising across continents and aggregating taxonomic resolution

### 2021-07-13 Paper/Submission_new/Latex_Supplementary_File.tex
13.07: Update revisions

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{float}
\usepackage[table,xcdraw]{xcolor}
\usepackage{color, colortbl}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{rotating}
\usepackage{caption}
\usepackage{authblk}
\usepackage{csquotes}
\graphicspath{{Figures/}}
\usepackage{setspace}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{array}
\usepackage{lscape}
\usepackage{longtable}
\usepackage{etoolbox}
\usepackage{hhline}
\usepackage{lmodern}

@@

    \begin{longtable}{m{1.9cm}|m{3.4cm}|m{3.3cm}|m{2.7cm}|m{3cm}|m{3.6cm}|m{2.5cm}}
        \caption{Comparison of trait definitions between invertebrate trait databases. Only traits that are differently described across databases are listed. The definition is quoted if it enables differences to be identified, otherwise the differences are described. The hyphen indicates a missing trait. Reproduction was captured in multiple grouping features per database. Hence, differences for reproduction have been described in the paper. Body form traits are not different between databases, except that the Vieira database contains the trait Bluff (blocky) which does not appear in the other databases.}

@@

        Distinguishes between active and passive &

@@

            \item bi/multivoltine

@@

        Distinguishes swimmer and skater & 

@@

            \item Differentiates also the trait interstitial (endobenthic)

@@

        Database contains traits crawler, 
        sprawler, climber and clinger. &

@@

        Distinguishes temporarily and permanently attached & 

@@

        Distinguishes spiracular gills, plastron, atmospheric breathers and plant breathers &
        Plastron and spiracle (termed aerial) occur as separate and combined traits. Contains also traits: air (plants), atmospheric, and functional spiracles &
        Distinguishes plastron and spiracle (termed aerial) \\

@@

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=18.5cm, height=16cm]{boxplot_scores_combined_REMAIN_SI.png}
    \caption{RDA of traits constrained by electric conductivity for the data aggregated with \textit{direct\_agg \textsubscript{mean}}, \textit{stepwise\_agg \textsubscript{mean}}, and \textit{weighted\_agg}. Shown are boxplots of the site scores along the conductivity axis. The rug on the right side of each plot indicates species scores of the traits on the conductivity axis. For more details see Figure 6. %\ref{fig:boxplot_scores_on_constrained_axis}. 
    Abbreviations: lcd, life cycle duration; nr.cy, potential number of cycles per year.}
    \label{fig:boxplots_scores_on_constrained_axis_REMAIN}
\end{figure}

>>>>

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{float}
\usepackage[table,xcdraw]{xcolor}
\usepackage{color, colortbl}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{rotating}
\usepackage{caption}
\usepackage{authblk}
\usepackage{csquotes}
\graphicspath{{Figures/}}
\usepackage{setspace}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{array}
\usepackage{lscape}
\usepackage{longtable}
\usepackage{etoolbox}
\usepackage{hhline}
\usepackage{lmodern}
\usepackage[
backend = biber,
natbib,
citestyle = authoryear,
bibstyle = apa,
maxcitenames = 2, 
maxbibnames = 99, 
uniquename=false,
uniquelist=false
%sorting = none % sort by name year title
]{biblatex}
\addbibresource{Ref_invertebrate_DB.bib}

@@

    \begin{longtable}{m{2cm}|m{3.25cm}|m{3.2cm}|m{3.1cm}|m{2.9cm}|m{3.4cm}|m{2.4cm}}
        \caption{Comparison of trait definition differences between invertebrate trait databases. Only traits that are differently described across databases are listed. The definition is quoted if it enables differences to be identified, otherwise the differences are described. Bullet points indicate when several traits have been used to describe a morphological or behavioural property. The hyphen indicates a missing trait. Reproduction was captured in multiple grouping features per database. Hence, differences for reproduction have been described in the paper. Body form traits are not different between databases, except that the Vieira database contains the trait Bluff (blocky) which does not appear in the other databases.}

@@

            \item Active filter feeders
            \item Passive filter feeders

@@

            \item bi - or multivoltine

@@

            \item Swimmer
            \item Skater

@@

            \item Interstitial (endobenthic)

@@

            \item Crawler
            \item Sprawler
            \item Climber
            \item Clinger

@@

            \item Temporarily attached
            \item Permanently attached

@@

            \item Spiracular gills
            \item Plastron
            \item Atmosph. breathers
            \item Plant breathers
            \item Plastron and spiracle (termed aerial) occur as separate and combined traits
            \item Air (plants)
            \item Atmospheric
            \item Functional spiracles
            \item Plastron
            \item Spiracle (termed aerial)

@@

\paragraph*{Discussion: Comparison of aggregation methods with varying phylogenies and trait variability}
\hfill
\\

We expected that, in addition to averaging measures, different weighting approaches to aggregation (i.e., equal weight to each taxon, equal weight to each genus, or weighted by number of species) would affect affinities for families with varying phylogenetic structure. However, we found through simulations that trait variability had a greater effect on the range of affinities (i.e., greater ranges of affinities at higher levels of variance) and on the differences in affinities between weighting approaches. In the simulation, taxonomic structural unevenness appeared to produce some variation in affinities among weighting approaches, especially for a family where one genus had a much larger number of species than its other genera. The comparison between aggregated and assigned traits showed, that the number of differing cases and the mean absolute differences in trait affinities, as well as the distributions of absolute trait affinity differences to assigned traits, were similar across the mean aggregation and across the median aggregation methods, which suggests a small influence of the weighting approach on the aggregation outcomes. The minor impact of the weighting approaches on trait aggregation might be explained by the fact that a considerable portion of taxa had low numbers of genera or species. Of the taxa that were compared from the North American trait dataset, 14 \% were identified at family and 62 \% at genus level, 52 \% comprised five or fewer genera, and 13 \% contained just one genus (Figure \ref{fig:tax_hierarchy_NOA}). In the Australian dataset, 21 \% of the compared taxa were identified at family, 40 \% at genus, and 39 \% at species level, 68 \% of the taxa contained five or fewer genera, and 40 \% just one genus (Figure \ref{fig:tax_hierarchy_AUS}). Hence, these results could change when more species-level trait information becomes available. Our findings show that the choice of aggregation method may matter less when phylogenetic structure is fairly even and traits are less variable, but the \textit{stepwise\_agg\textsubscript{median}} method tended to produce the widest range in affinities, especially with high taxonomic unevenness and high trait variability.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=18.5cm, height=16cm]{boxplot_scores_combined_REMAIN_SI.png}
    \caption{RDA of traits constrained by electric conductivity for the data aggregated with \textit{direct\_agg \textsubscript{mean}}, \textit{stepwise\_agg \textsubscript{mean}}, and \textit{weighted\_agg}. Shown are boxplots of the site scores along the conductivity axis. The rug on the right side of each plot indicates species scores of the traits on the conductivity axis. For more details see Figure 6. %\ref{fig:boxplot_scores_on_constrained_axis}. 
    Abbreviations: lcd, life cycle duration; nr.cy, potential number of cycles per year.}
    \label{fig:boxplots_scores_on_constrained_axis_REMAIN}
\end{figure}

\subsubsection*{Fourth corner analysis}

We selected the ecologically relevant traits identified by \citet{szocs_effects_2014} that increased (shredder, life cycle duration $ > 1 $ year, gills, bi/multivoltine, ovoviviparity) or decreased (univoltine, eggs in clutches/cemented or fixed, life cycle duration $ <= 1$ year) with increasing salinity and used these in the fourth corner analysis. The selected traits have also reacted to salinity in other studies (but see the discussion in \citet{szocs_effects_2014}). 
For some sites, environmental and abundance data have been sampled multiple times per year. \citet{szocs_effects_2014} used the full data and did not consider repeated sampling. We applied two analysis: one where we did not consider multiple samplings. And second where we selected, for the sites with multiple samplings, only those sampling events with the highest macroinvertebrate abundances per year and accordingly only used the conductivity data for these sampling events (this was not done for the RDA approach).Fourth corner analysis was carried out in R with the \textit{fourthcorner()} function from the ade4 package.
With the fourth corner analysis fewer trait-conductivity relationships than with the RDA approach were detected. With the original dataset significant relationships to conductivity were found for the traits shredder and ovoviviparity. In three out of the six aggregated and harmonised datasets (\textit{direct\_agg\textsubscript{mean}}, \textit{weighted\_agg}, harmonised (not aggregated)) four to five significant relationships (for the traits shredder, univoltinism, bi/multivoltine, ovoviviparity, and aquatic oviposition) were detected. In the remaining three datasets (\textit{stepwise\_agg\textsubscript{median}}, \textit{stepwise\_agg\textsubscript{mean}}, and \textit{direct\_agg\textsubscript{median}}) two significant relationships for the traits univoltinism and bi/multivoltinism were detected. Results were only slightly different when not considering multiple samplings (original dataset: one significant trait-conductivity relationship, others three to four). For the significant trait-conductivity relationships, the classification into characteristic traits for downstream and upstream sites was consistent with the results of the RDA analysis.

\subsection*{Effect of harmonisation and trait aggregation on functional diversity metrics}

\begin{figure}[ht]
    \centering
    \includegraphics[width=17cm, height=12cm]{Correlation_fd_metrics_FRic.png}
    \caption{Fitted regressions between residuals of FRic with traits used in the original study and with harmonised and aggregated traits. Dots refer to each site-year combination. Sites have been sampled maximum three times in three years. For further details please refer to sections "Effects of harmonisation and trait aggregation on inferences regarding trait-environment relationships" and the original study.}
    \label{fig:FRic}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=17cm, height=12cm]{Correlation_fd_metrics_FDiv.png}
    \caption{Fitted regressions between residuals of FDiv with traits used in the original study and with harmonised and aggregated traits. Dots refer to each site-year combination site-year combination. Sites have been sampled maximum three times in three years. Some sites were sampled multiple times within a year. In that case the highest sampling event with the highest total invertebrate abundance was taken (this explains the discrepancy to the original study, that did not consider multiple samplings). For further details please refer to sections "Effects of harmonisation and trait aggregation on inferences regarding trait-environment relationships" and the original study.}
    \label{fig:FDiv}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=17cm, height=12cm]{Correlation_fd_metrics_FEve.png}
    \caption{Fitted regressions between residuals of FEve with traits used in the original study and with harmonised and aggregated traits. Dots refer to each site-year combination site-year combination. Sites have been sampled maximum three times in three years. Some sites were sampled multiple times within a year. In that case the highest sampling event with the highest total invertebrate abundance was taken (this explains the discrepancy to the original study, that did not consider multiple samplings). For further details please refer to sections "Effects of harmonisation and trait aggregation on inferences regarding trait-environment relationships" and the original study.}
    \label{fig:FEve}
\end{figure}
